---
title: "Practicum 3"
author: "Jenny Balmagia, Anthony Luna, Hope Cupples"
date: "May 27, 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DeclareDesign)
library(truncnorm) #for truncated distribution
library(knitr)
library(ggplot2)
library(kableExtra)
```

##declare_population()

```{r population, echo=TRUE}
set.seed(228)
population <- declare_population(
  households = add_level(N=1000000,
    kwh = rnorm(n=N, mean = 550, sd = 75),
    u=rnorm(n=N, mean= 0.75, sd= 0.1)) #25% reduction w/ 10% sd due to retrofit 
)

pop <- population()

#1,000,000 is an estimate of the number of households in the bay area, kwh is a distribution of household energy usage per month (Kwh) and u is the estimated percent change in energy usage from baseline due to retrofits - not including the pledge, which is the treatment. 


te <- 0.97 #Taking the pledge will result in an additional 3% reduction in energy usage per month
samp_size <- 5000 #half of the households recieving retrofits will also be asked to take the pledge

potential_outcomes <- 
  declare_potential_outcomes(
    Y_D_0=kwh*u,
    Y_D_1= kwh*u*te)#Multiplied TE since it is also a % reduction, not a fixed number of KwH

po <- potential_outcomes(pop)
kable(po[1:5,], digits=1)

sampling <- declare_sampling(n=samp_size)
sam <- sampling(po)
kable(sam[1:5,c(1:2,4:6)], row.names = FALSE,
      digits = 1)

assigning <- declare_assignment(prob = 0.5,
                  assignment_variable="D")
assigned <- assigning(sam)
kable(assigned[1:5,c(1:2,4:5,7:8)], 
      digits = 1)

ggplot(data=assigned, aes(x=as.factor(D), y=kwh)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")

revealing <- declare_reveal(assignment_variables=D)

estimand <- declare_estimand(ATE = te)
estimand(po)

dim <- declare_estimator(Y ~ D, estimand = estimand,  
          model =  difference_in_means, label = "DIM")

did <- declare_estimator(Y - kwh ~ D, 
                         estimand = estimand,  
          model =  difference_in_means, label = "DID")

design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + did

diagnosis <- diagnose_design(design, sims=100)
diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
  kable()
```



##declare_sampling()

Next, we want to select the sample size. Let's start with 100 villages (recall that the actual study used 120 villages)

```{r sample, echo=TRUE}
sampling <- declare_sampling(n=5000)
sam <- sampling(po)
kable(sam[1:5,c(1:2,4:6)], row.names = FALSE,
      digits = 1)
```

##declare_assignment()

This step declares the random assignment process. There are many complexities, but let's stick to *complete assignment* of exactly half the units at this stage.

```{r assign, echo=TRUE}
assigning <- declare_assignment(m = nrow(sam)/2,
                  assignment_variable="D")
assigned <- assigning(sam)
kable(assigned[1:5,c(1:2,4:5,7:8)], 
      digits = 1)

#PH: alright, now we're using DeclareDesign to randomly assign villages to treatment and control with a probability of 0.5. The variable 'm' denotes the number of units that declare_assignment() should place in each treatment condition, which we are asking to be equal to half of the sample per treatment condition (nrow(sam)/2).

```


## Assessing balance

At this stage, it's possible to look at balance in the baseline tree cover characteristics, since random assignment has occured.

```{r violin, echo=FALSE, fig.height=6}
ggplot(data=assigned, aes(x=as.factor(D), y=wqi)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")

#PH: recall that random assignment should, in expectation, produce treatment and control groups that are statistically-identical on all observed and unobserved features.
```

##declare_reveal()

This step declares how the potential outcomes are revealed by the random assignment

```{r reveal, echo=TRUE}
revealing <- declare_reveal(assignment_variables=D)
#PH: this part of DeclareDesign essentially is telling R which PO to reveal for each unit as a function of its treatment assignment. Recall that we calculated the full schedule of potential outcomes for each unit above.
```

##declare_estimand()

At this stage, we specify our target *estimand*, which is the quantity that we are trying to recover when estimating impact. Recall that we set this value to **5.5** in line with Jayachandran et al. (2017).

```{r estimand, echo=TRUE}
estimand <- declare_estimand(ATE = -0.19)
estimand(po)
```


##declare_estimator()

Next, we declare the estimators we use for recovering the estimand. While there are many advanced estimators, we'll focus on the two core experimental estimators:
1. difference-in-means
2. difference-in-differences

```{r estimator, echo=TRUE}
dim <- declare_estimator(Y ~ D, estimand = estimand,  
          model =  difference_in_means, label = "DIM") #PH: Difference-in-means (Mean of treatment group - mean of control group)

did <- declare_estimator(Y - wqi ~ D, 
                         estimand = estimand,  
          model =  difference_in_means, label = "DID") #PH: Difference-in-differences ([Mean of treatment group @ endline - Mean of treatment group @ baseline] - [Mean of control group @ endline - mean of control group @ baseline])
```


##declare_design()

This function brings all of the parts of the process together in a single design and allows for each part of the design to be simulated repeatedly.

```{r design, echo=TRUE}
design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + dim + did
```


##diagnose_design()

At this stage, we can calculate various features of the design that we have specified

```{r diagnosis, cache=TRUE}
diagnosis <- diagnose_design(design, sims=500)
diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
  kable()
```


## Looking under the hood, DIM

```{r underhood-dim, height=6}
sim.out <- diagnosis$simulations
hist(sim.out$estimate[sim.out$estimator_label=="DIM"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-2,2), cex=24)
abline(v=-0.19, lwd=3, col="red", add=T)
```

## Looking under the hood, DID

```{r underhood-did, height=6}
sim.out <- diagnosis$simulations
hist(sim.out$estimate[sim.out$estimator_label=="DID"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-2,2), cex=24)
abline(v=-0.19, lwd=3, col="red")

#PH: note that we get more power using the difference in differences estimation. This is because diff-in-diff increases the precision of our estimate of the ATE.
```

